{How do we copy and paste?\\The semantic drift of quotations in blogspace}

### 2014-12-12 brains-introduction.tex
introduction / hypotheses mods

The understanding of knowledge transmission mechanisms has led to a sizable literature in the recent past, spanning over numerous research fields ranging from cultural anthropology to social network analysis and complex systems modeling, from social cognition to data mining. These works are diversely labeled as studies on ``opinion dynamics'', ``cultural evolution'', or ``information diffusion'' and, for the most part, 
%Broadly speaking, this type of research program is bound to
 investigate phenomena pertaining to both cognitive science and social science, %and aims to understand the processing, transmission, and evolution of information 
 both at the individual and social levels.  Broadly speaking, we may distinguish two main research streams, depending on whether the focus lies on cognitive processes or on social dynamics.

%Broadly speaking, this type of research program is bound to investigate phenomena pertaining to both cognitive science and social science, and aims to understand the processing, transmission, and evolution of information both at the individual and social levels.  The relevant literature features two main streams, depending on whether the focus lies on social dynamics or cognitive processes.

A first stream of research is largely structured around cultural anthropology and essentially addresses cultural similarity, diversity and its evolution. It features several theories mixing social and individual cognition including, to cite a few, the debate around the ``memetic'' program initiated by \citet{Dawkins76} (for which the collection of works by \citet{Aunger00} provides a solid overview), the development of evolutionary models of norms (see for instance \citet{Ehrlich05}, following the seminal work of \citet{Boyd85}); or the ``cultural epidemiology'' program initiated by \citet{sper:expl}, which %recently attracted significant attention, and
% focuses on the notion of \emph{representation}, 
links %the cognitive science 
the concept of \emph{mental representation} to the concept of \emph{public representation} (the latter being the former's counterpart outside the brain, i.e. in cultural artifacts: texts, utterances, etc.).\footnote{\citet{sper:expl} emphasizes this distinction in his seminal work:
\begin{quotation}
A representation may exist inside its user: it is then a \emph{mental representation}, such as a memory, a belief, or an intention.
The producer and the user of a mental representation are one and the same person.
A representation may also exist in the environment of its user, as is the case, for instance, of the text you are presently reading: it is then a \emph{public representation}.
\end{quotation}}

Here, one of the core claims consists in observing that not all knowledge is equally fit for being reproduced, although the various approaches have a different take on how exactly this notion of fitness should be operationalized. Sperber's cultural epidemiology opposes Dawkins' memetics by insisting that representations are not being replicated through a high-fidelity copy process, but are being interpreted and produced anew, and are thus greatly subject to change.
Cultural epidemiology postulates this type of conceptual evolution and suggests that it can be appraised through the notion of ``cultural attractor'', seen as the attraction domain of an underlying socio-semantic dynamical system.\footnote{Works such as \citet{Atran03} argue that this approach is anthropologically better suited than memetics, and the main issues in this debate are further detailed by \citet{Kuper00} and \citet{Bloch00}.}
Despite some recent modeling attempts (for instance \citet{Claidiere07}), the development of quantitative measurements focused on the key notion of cultural attractor has remained a relatively hard task and, to our knowledge, this hypothesis has not yet been empirically analyzed in an extensive manner.

@@

These studies are relatively remote from anthropology and cognition, and, for a significant part, rely only marginally on specific social science theories; they are rather at the interface between data mining, complex systems and quantitative sociology (first and foremost social network analysis). They nonetheless show us the interest of using these records --~given the already significant and rapidly growing importance of our online interactions~--
%and, %more broadly, large textual corpora \cam{google n-grams??} 
to dramatically improve the empirical understanding of individual-level processes of cultural evolution.

\bigskip
Stepping back, we observe a gap between, on one side, empirical studies of the diffusion dynamics in a social system and, on the other side, more theoretical works  focused on knowledge transformation processes. %--- these latter studies being either strongly normative, or with results difficult to articulate with realistic cultural epidemiology models.
Our research lies at the intersection of these two programs, aiming to shed light on microscopic information transformation processes by leveraging the empirical wealth of (in vivo) social diffusion phenomena.   We hope to describe some reformulation processes within a large distributed social system such as blogspace. More precisely, we aim to empirically describe how some specific type of public representation may possibly be altered by bloggers when they try to reproduce that representation.

>>>>

The understanding of knowledge transmission mechanisms has led to a sizable literature in the recent past, spanning over numerous research fields ranging from cultural anthropology to social network analysis and complex systems modeling; from social cognition to data mining. These works are diversely labeled as studies on ``opinion dynamics'', ``cultural evolution'', or ``information diffusion'' and, for the most part, investigate phenomena pertaining to both cognitive science and social science, both at the individual and social levels.

Broadly speaking, we may distinguish two main research streams, depending on whether the focus lies on cognitive processes or on social dynamics. A first stream of research is largely structured around cultural anthropology and essentially addresses cultural similarity, diversity and its evolution. It features several theories mixing social and individual cognition including, to cite a few, the debated ``memetic'' program initiated by \citet{Dawkins76} (for which the collection of works by \citet{Aunger00} provides a solid overview), the development of evolutionary models of norms (see for instance \citet{Ehrlich05}) following the seminal work of \citet{Boyd85}); or the ``cultural epidemiology'' program proposed by \citet{sper:expl}, which links the concept of mental representation to the concept of public representation (the latter being the former's counterpart outside the brain, \hbox{i.e.} in all sorts of cultural artifacts: texts, utterances, etc.).
\footnote{\citet{sper:expl} emphasizes this distinction in his seminal work: \begin{quotation}A representation may exist inside its user: it is then a \emph{mental representation}, such as a memory, a belief, or an intention. The producer and the user of a mental representation are one and the same person. A representation may also exist in the environment of its user, as is the case, for instance, of the text you are presently reading: it is then a \emph{public representation}.\end{quotation}}

Here, one of the core claims consists in emphasizing that not all knowledge is equally fit for being reproduced, although the various approaches have a different take on how exactly this notion of fitness should be operationalized. Sperber's cultural epidemiology opposes Dawkins' memetics by insisting that representations are not being replicated through a high-fidelity copy process, but are being interpreted and produced anew, and are thus greatly subject to change.
Cultural epidemiology postulates that this conceptual evolution can be appraised through the notion of ``cultural attractor'', seen as the attraction domain of an underlying socio-semantic dynamical system.\footnote{Works such as \citet{Atran03} argue that this approach is anthropologically better suited than memetics, and some of the main issues in this debate are further detailed by \citet{Kuper00} and \citet{Bloch00}.}
Despite some recent modeling attempts (\hbox{e.g.} \citet{Claidiere07}), the development of quantitative measurements relying on the concept of cultural attractors has remained a relatively hard task and, to our knowledge, this hypothesis has not yet been empirically analyzed in an extensive manner.

@@

These studies are relatively remote from anthropology and cognition, and, for a significant part, rely only marginally on specific social science theories; they are rather at the interface between data mining, complex systems and quantitative sociology (first and foremost social network analysis). They nonetheless show us the interest of using these rapidly growing records %--~given the already significant and rapidly growing importance of our online interactions~--
%and, %more broadly, large textual corpora \cam{google n-grams??} 
in radically improving the empirical understanding of (individual-level) cultural evolution processes.

\bigskip
Stepping back, we thus observe a gap between, on one side, empirical studies of diffusion dynamics in social systems and, on the other side, more theoretical works  focused on knowledge transformation processes. %--- these latter studies being either strongly normative, or with results difficult to articulate with realistic cultural epidemiology models.
Our research lies at the intersection of these two programs, aiming to shed light on micro-level information transformation by leveraging the empirical wealth of (in vivo) social diffusion phenomena. More precisely, we hope to describe reformulation processes within a large distributed system such as blogspace; showing how some specific types and features of public representations may be altered by bloggers when they freely reproduce that representation.
